Investigating soundscapes perception through acoustic scenes simulation.

### 2017-12-20 lafaySimulatedScenePerception.tex
vtest ref

\documentclass[12pt]{elsarticle}
\usepackage{url}
\usepackage[utf8]{inputenc}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{caption}
\usepackage{sectsty}
\usepackage{amsmath}
\usepackage{color}

@@

\newcommand{\gl}[1]{\textcolor{magenta}{GL : #1}}
\newcommand{\mr}[1]{\textcolor{green}{MR : #1}}
\newcommand{\ml}[1]{\textcolor{red}{ML : #1}}
\newcommand{\jfp}[1]{\textcolor{blue}{JFP : #1}}
\newcommand{\ie}{\emph{i.\,e.}}
\newcommand{\Ie}{\emph{I.\,e.}}
\newcommand{\cf}{cf.}
\newcommand{\Cf}{Cf.}
\newcommand{\vs}{\emph{vs.}}
\newcommand{\Vs}{\emph{Vs.}}
\newcommand{\al}{\emph{et~al.}}
\newcommand{\eg}{\emph{e.\,g.}}
\newcommand{\Eg}{\emph{E.\,g.}}
\newcommand{\etc}{\emph{etc.}}
\newcommand{\myfloatalign}{\centering}

\journal{Behavior Research Methods}

@@

    \author{G. Lafay$^{1)}$, M. Rossignol$^{2)}$, N. Misdariis$^{2)}$, M. Lagrange$^{1)}$, J-F. Petiot$^{1)}$}

@@

  \begin{abstract}
    This paper introduces a new experimental protocol to study mental representations of urban soundscapes through a simulation process. Subjects are asked to create a full soundscape by means of a dedicated software tool, coupled with a structured sound data set. This paradigm is used to characterize urban sound environment representations by analyzing the sound classes that were used to simulate the auditory scenes. A rating experiment of the soundscape pleasantness using a 7-point bipolar semantic scale is conducted to further refine the analysis of the simulated urban acoustic scenes. Results show that 1) a semantic characterization in terms of presence~/ absence of sound sources is an effective way to characterize urban soundscapes pleasantness, and 2) physical descriptors computed for specific sound sources better characterize the appraisal than global descriptors.
  \end{abstract}

  \begin{keyword}
  %% keywords here, in the form: keyword \sep keyword
 cognitive psychology \sep soundscape perception \sep soundscape simulator
  %% PACS codes here, in the form: \PACS code \sep code

@@

One of the main goal of soundscape studies is to identify which components of the soundscape influence human perception \cite{aletta2016soundscape}. For example, in order to study the link between an environment (urban sound scene) and the induced human sensation (calmness), one mean is to objectify the mental representation of a specific sound scene (a calm urban sound scene). Several means have been considered in the literature to perform this task.

@@

First, a subject can be asked to assess a given sound following a given perceptual scale (\eg~calmness, pleasantness) \cite{axelsson2005soundscape,davies2013perception,cain2013development}. The amount of information that can be gathered strongly depends on the very nature of the available stimuli, be they recorded sound scenes as part of a within laboratory experiment, or \emph{in situ}. That being said, those recorded stimuli  can be analyzed, what allows the experimenter to gather a coarse
grain physical description of them.

@@

Second, a subject can be asked to describe a given sound environment \cite{guastavino2006ideal, dubois2006cognitive}. A large amount of quantitative and semantic information is then collected about the subject's representation of this type of sound environment. Unfortunately, without any reference to sound data, this representation can hardly be characterized physically.

@@

We propose in this paper to consider the use of a soundscape simulator that the subject can manage to objectify his/her representation of a given sound environment. We believe that the use of such a device allows us to gain the benefits of the two above cited approaches. As the subject is asked to produces audio data (the signal of the simulated scene), it allows the experimenter to study a precise modal version of the mental representation of the subject that is both characterized semantically and physically, respectively in terms of the sound sources and their levels.

@@

@@

For several reasons that are detailed in this paper, the use of a soundscape simulator such as the one proposed in this article can lead to interesting outcomes as it allows the experimenter to separately study  the influence of the sound sources that compose a sound environment of interest. With such a material, not only the type of sound sources is indeed available for study, but also the exact level and audio waveform for each source together with the structural characteristics of the scene, that is the temporal distribution of the sound events.

@@

To the best of our knowledge, only Bruce~\cite{bruce2009development,bruce2014effects} considered the use of a simulator to question soundscape perception. They propose a tool that allows the user to modify a given soundscape by adding or removing specific sound sources, by changing the acoustic level of the sources as well as their spatial location. The authors shows that the addition or removal of the sources globally follows more social or semantic matters than their acoustical characteristics. A lack of diversity in terms of sound sources is nonetheless mentioned by the authors that limit the strength of the outcomes given in this study.

@@

The remaining of the paper is organized as follows: the soundscape simulator \emph{simScene} is introduced in Section \ref{sec:simulator}. The experiments  1.a \emph{simulation} and  1.b \emph{pleasantness evaluation} are presented in Section \ref{sec:simulation}. Experiment 2 \emph{pleasantness evaluation after modification of the scenes} is presented in Section \ref{sec:modification}. Conclusions and discussion about future work follow in Section \ref{sec:conclusion}.

@@

\emph{Simscene} is an online digital audio tool whose first version has been developed as part of the HOULE project\footnote{\emph{Projet HOULE} : \url{houle.ircam.fr}}. It has been designed to run on the popular web browsers \emph{Chrome} and \emph{Firefox}. It is fully written in javascript using the \emph{angular.js} library\footnote{\emph{angular.js} : \url{angularjs.org}} and the \emph{web-audio} standard that allows the manipulation of digital audio data within the browser\footnote{\emph{web-audio}: \url{www.w3.org/TR/webaudio}}. The interface for selecting the sound sources (\cf~Section~\ref{sec:simscene_ssf}) uses the popular \emph{D3.js} \cite{d32011} visualization library.

@@

\emph{Simscene} is designed as a simplified audio sequencer, with sequencing parameters specifically chosen for the generation of realistic soundscapes. To do so, the user first selects a sound source using a non verbal selection interface presented in Section~\ref{sec:simscene_ssf}. A track is then created for this sound source within the simulator interface. The user can then manipulate some parameters detailed in Section~\ref{sec:simscene_parametre} to control the time and magnitude distribution of the occurrences of the sound sources. Text fields are also available for the user to 1) name each track, 2) name the entire scene, 3) to provide some comments about the simulated scene.

@@

@@

In order to provide the user with a sound database that is well organized and covers as much as possible the variety of sound sources that are present in urban areas, a typology of urban sounds is first defined.

@@

The literature usually distinguishes between:

\begin{itemize}
\item {a sound event}: an isolated sound, sparsely distributed whose acoustical characteristics may change with respect to time
\item {a sound texture}: an isolated sound of long duration whose acoustical characteristics are stable with respect to time \cite{saint1995classification}.
\end{itemize}

@@

The simulation tool \emph{Simscene} follows this distinction and considers two distinct sound databases: one with the classes of events only and the other with the classes of textures only. These two types of sound classes also have specific scheduling procedures during the simulation process.

@@

According to the Auditory Scene Analysis (ASA) theory, the brain processes separately events emerging from the different sources of a sound environment \cite{bregman1994auditory}. Several auditory streams are created, one for each sequence of events emitted by the same source \cite{carlyon2004brain}. In this case, the source is a perceptual entity as events that are not isolated are grouped into a stream.

@@

Maffiolo \cite{maffiolo_caracterisation_1999} distinguishes two distinct categorization processes, that differentiate themselves with the capacity of the listener to identify sound events. She shows that those processes leads to two abstract cognitive category respectively termed "the event sequences" and "the amorphous sequences".

@@

The event sequences are composed of salient events that can easily be recognized such as \emph{car start} or \emph{male speech}. They arise from a descriptive analysis based on the identification of the sound sources. On the contrary, the amorphous sequences are sound environments where distinct events can hardly be found, such as \emph{traffic hubhub}. They result from a holistic analysis based on global acoustic features.

@@

\begin{figure}[t]
        \myfloatalign
        \includegraphics[width=.8\linewidth]{gfx/3.pdf}
       \caption{Hierarchical organisation of the isolated sounds used in the simulation.\gl{TODO : translate and remove subject box}}\label{fig:orgDb}
\end{figure}

@@

We term "sample", a recording of an isolated sound, be it an event or a texture. Each sound class is implemented as a collection of samples perceived equivalently.

@@

The sound classes are organized hierarchically (\cf~Figure~\ref{fig:orgDb}) according to a structure that is close to the one of the vertical axis of the categorical organization proposed by E. Rosch \cite{rosch1978cognition}. The lower the level of abstraction, the more precise the description of the class and the more perceptually similar the sound sources. For classes with high level of abstraction that have sub classes, their collection of samples is the union of the collections of the sub classes.

@@

Accounting for the perceptual matters detailed previously, two taxonomies are built, one for event sounds and one for texture sounds. Four levels of abstraction are considered from the more generic one (level 0), to the more specific classes (level 2), leading to a taxonomy close to the one used in \cite{Salamon14}. Only three levels of abstraction are considered for the texture sounds.

@@

All samples are normalized to the same $RMS$ level of $-12$ $dB$ FS, \ie relative to Full Scale. In our case, the full scale level is set arbitrarily to 1 Volt.

@@

By a voluntary design choice, the simulation tool does not allow the user to interact and control directly a specific sample. Interaction is done at the track level, a track being a sequence of samples. Several parameters are available to the subject to control the track:

@@

\begin{itemize}
\item \emph{sound level} ($dB$): for each sample, the sound levels are randomly sampled according to a normal distribution parametrized by the subject in terms of mean value and variance;
\item \emph{inter-onset spacing} (second): for the event's track only, and for each sound event sample, the inter-onset spacings are randomly sampled according to a normal distribution parametrized by the subject in terms of mean value and variance;
\item \emph{start and end time} (second): the subject set the first and last occurrence of a sample of the selected class.
\end{itemize}

@@

\begin{itemize}
\item \emph{event's fades} (seconds): for the event's track only, the subject can set a fade in / fade out duration applied to each sample;
\item \emph{global fades} (seconds): the subject can set global fade in / fade out duration applied to the entire track.
\end{itemize}

@@

\begin{figure}[!tp]
		\myfloatalign
        \subfloat[]
        {\includegraphics[width=.8\linewidth]{gfx/SSF}\label{fig:ssf}} \par
        \subfloat[]
        {\includegraphics[width=\linewidth]{gfx/simscene}\label{fig:simscene}}
       \caption{Graphical interfaces for the selection of sound classes (a) and their sequencing (b) of the \emph{simScene} tool.}
\end{figure}

@@

Once the typology and the set of sounds are available, an important design issue is the need of a suitable way to display the sound dataset to the user. Most browsing tools are based on keyword indexing. However, for sensory experiment that study the objectivation of the mental representation of the subject, this mean is problematic as the availability of a verbal description of the sound can influence the choice of the subject and potentially induce biases in the analysis conclusions. Indeed, the subject may select \textit{a priori} sounds referenced as belonging to \emph{park} to build a \emph{calm} soundscape.

@@

Therefore, the selection interface considered in this study is text free and composed of spatially organized circles on a plane, each representing one leaf class.

@@

@@

As shown on Figure~\ref{fig:simscene}, the simulation interface displays a schematic of the scene under creation. Each track is represented as an horizontal band with a temporal axis. Each sample of this track is displayed by a rectangle whose height is proportional to the amplitude of the sample. For event's tracks, the horizontal spacing between those rectangles is a function of the time delay between their onsets. For tracks of texture, a unique rectangle is displayed as this kind of sounds does not allow spacing with silence. As actual amplitude and spacing values are drawn from random variables, each time the subject changes the value of a parameter, the location and height of the rectangles are updated to reflect the changes in the sequencing of the samples. The subject can listen to the resulting sound scene at any time.

@@

\begin{figure}[t]
        \includegraphics[width=\linewidth]{gfx/5.pdf}
        \caption{Experimental protocol of the simulation experiment (1.a) and the pleasantness evaluation experiment (1.b).\gl{TODO : translate}}\label{fig:xp1_2}
\end{figure}

% L'objectif est d'étudier les influences spécifiques des différentes sources sonores qui composent les environnements urbains sur la perception de l'agrément, en utilisant la simulation. Pour ce faire, nous planifions nos deux premières expériences (\cf~Figure~\ref{fig:xp1_2}):

Experiment 1 aims at using a simulation paradigm to investigate the specific influences of the various sound sources constituting urban soundscapes on the perceived pleasantness. For that, the two first experiments are planned as follows (\cf~Figure~\ref{fig:xp1_2}):

% \begin{itemize}
% \item \emph{expérience 1.a de simulation}: au cours de cette expérience, les sujets doivent simuler des environnements sonores urbains, en utilisant l'outil de simulation \emph{Simscene} (\cf~Section~\ref{sec:simscene}). Chacun compose deux environnements sonores, le premier idéal/agréable et le deuxième non-idéal/désagréable.
% \item \emph{expérience 1.b d'évaluation}: à l'issue de la simulation, nous n'avons, de fait, qu'une connaissance binaire des propriétés affectives des scènes simulées: idéale (i) et non-idéale (ni). Cette seconde étape a pour but d'affiner notre connaissance sur l'agrément. Pour ce faire, nous demandons à un deuxième groupe de sujets d'évaluer, à partir d'une échelle sémantique, l'agrément de chacune des scènes simulées. L'expérience d'évaluation sert deux buts:
% \begin{enumerate}
% \item évaluer l'influence respective des différentes sources sur l'agrément, pour chaque type d'environnement (i ou ni);
% \item détecter la présence de cas extrêmes ou ambigus (\emph{outlier}) dans les scènes simulées. Pour le reste de notre étude, les qualités hédoniques imposées (i et ni) servent de référence, de vérité terrain. Il nous faut donc garantir qu'il n'y ait pas d’ambiguïté entre les cas extrêmes des i- et ni-scènes, \ie~que la note d'agrément la plus basse des i-scènes reste supérieure à la note la plus haute des ni-scènes.
% \end{enumerate}
% \end{itemize}

\begin{itemize}
\item \emph{experiment 1.a (simulation)}: during this experiment, participants are asked to simulate urban sound environments, using the simulation tool \emph{Simscene}, see Section~\ref{sec:simulator}. Each of them has to create two sound environments: one ideal / pleasant and the other one, non ideal / unpleasant.
\item \emph{experiment 2.a (evaluation)}: after the simulation phase, only a binary information on the pleasantness property of the respective scenes is available: respectively ideal (i) or non ideal (ni). Furthermore, this information is given by the creator of the scene. The second experimental step aims at investigating more deeply and more broadly our knowledge on the pleasantness of the simulated scenes. For that, a second group of participants is asked to evaluate the pleasantness of each scene, on a semantic scale. This experiment has two goals:
\begin{enumerate}
\item to evaluate more precisely the respective influence of the various sources composing the scenes on the pleasantness  (i or ni) thanks to a finer quantification of the pleasantness of the scene.
\item to detect the presence of \emph{outliers} or ambiguous scenes. Indeed, throughout our analyses, the predefined hedonic properties of the scenes (i and ni) are used as reference. We thus need to ensure beforehand that no ambiguity exists between extreme cases of i and ni scenes, \ie~that the less pleasant i-scene remains above the more pleasant ni-scene.
\end{enumerate}
\end{itemize}

% Notre analyse s'appuie sur les données produites par les deux expériences.

The data collected by these two experiments (1.a and 1.b) are analyzed conjointly.

@@

The design of this experiment has been validated with a pilot study described in \cite{lafay2014new}.

@@

% Les sujets doivent simuler deux environnements sonores urbains, chacune des scènes devant durer 1 minute. Ils doivent se conformer aux consignes suivantes:

The participants have to simulate two urban sound environments of one minute each. They follow the instructions below:

% \begin{itemize}
% \item première simulation: simuler un paysage sonore \textbf{urbain plausible} qui, selon vous, est idéal (où vous aimeriez vivre);
% \item deuxième simulation: simuler un paysage sonore \textbf{urbain plausible} qui, selon vous, est non-idéal (où vous n'aimeriez pas vivre).
% \end{itemize}

\begin{itemize}
\item  first simulation: create a \textbf{plausible urban soundscape} which is ideal, according to you (where you would like to live);
\item second simulation: create a \textbf{plausible urban soundscape} which is non ideal, according to you (where you would not like to live);
\end{itemize}

% Tous commencent par simuler l'environnement idéal. Ils ne prennent connaissance de la deuxième consigne qu'à la fin de la première simulation.

% Les sujets sont totalement libres dans le choix des sons, et des paramètres. Ils doivent cependant se soumettre à deux contraintes:

All the participants start by designing the ideal environment; they read the second set of instructions at the end of the first experiment. Participants are completely free of their choices concerning sounds and synthesis parameters. The created sound environments must nevertheless fulfill the two following constraints :

% \begin{itemize}
% \item le sujet doit prendre le point de vue d’un auditeur fixe;
% \item le paysage sonore doit être réaliste, au sens de physiquement plausible. Le sujet peut très bien réunir 10 chiens dans un paysage sonore. Il ne peut en introduire 1 qui aboierait toutes les 10 millisecondes.
% \end{itemize}

\begin{itemize}
\item the listening point of view is the one of a fixed listener;
\item the soundscape must be realistic, \ie~physically plausible. For instance, participants are free to insert ten dogs in the soundscape but they cannot insert one dog barking every 10 milliseconds.
\end{itemize}

% Ces contraintes font partie de la consigne. Aucun contrôle n'est fait \emph{a priori} dans l'interface de simulation.

These constraints are notified in the instructions; no control is done \emph{a priori} in the simulation interface.

% Chaque processus de simulation comprend deux parties :
%
% \begin{enumerate}
% \item la réalisation de la simulation: cette étape se décompose en trois temps:
% \begin{itemize}
% \item sélectionner les classes de sons
% \item nommer les classes de sons sélectionnées
% \item paramétrer les pistes (\cf~Section~\ref{sec:simscene_parametre}) relatives aux classes de sons sélectionnées
% \end{itemize}
% \item la production d'un commentaire libre du paysage sonore simulé
% \end{enumerate}

Each simulation process is decomposed in several steps:

\begin{enumerate}
\item the simulation: the user is asked to:
\begin{itemize}
  \item  select the class of sounds
	\item  give a name to the selected class of sounds
	\item  set the parameters of the tracks related to the selected class of sounds, see Section~\ref{sec:simscene_parametre}.
\end{itemize}
\item the feedback : writing of a free form comment about the composed soundscape
\end{enumerate}

>>>>

\documentclass[preprint,12pt]{elsarticle}
\usepackage{url}
\usepackage[utf8]{inputenc}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{caption}
\usepackage{sectsty}
\usepackage{amsmath}

@@

\newcommand{\gl}[1]{\textcolor{blue}{GL : #1}}
\newcommand{\nm}[1]{\textcolor{magenta}{#1}}
\newcommand{\ml}[1]{\textcolor{red}{#1}}
\newcommand{\jlc}[1]{\textcolor{green}{#1}}
\newcommand{\ie}{\emph{i.\,e.}}
\newcommand{\Ie}{\emph{I.\,e.}}
\newcommand{\cf}{cf.}
\newcommand{\Cf}{Cf.}
\newcommand{\vs}{\emph{vs.}}
\newcommand{\Vs}{\emph{Vs.}}
\newcommand{\al}{\emph{et~al.}}
\newcommand{\eg}{\emph{e.\,g.}}
\newcommand{\Eg}{\emph{E.\,g.}}
\newcommand{\etc}{\emph{etc.}}
\newcommand{\myfloatalign}{\centering}

\journal{Nuclear Physics}

@@

    \author{G. Lafay$^{1)}$, M. Rossignol$^{2)}$, N. Misdariis$^{2)}$, M. Lagrange$^{1)}$, J.F. Petiot$^{1)}$}

@@

  \begin{abstract}
    This paper introduces a new experimental protocol to study mental representations of urban soundscapes through a simulation process. Subjects are asked to create a full soundscape by means of a dedicated software tool, coupled with a structured sound data set. This paradigm is used to characterize urban sound environment representations by analyzing the sound classes that were used to simulate the auditory scenes. A rating experiment of the soundscape pleasantness using a 7-points Likert scale is conducted to further refine the analysis of the simulated urban acoustic scenes. Results show that 1) a semantic characterization in terms of presence~/ absence of sound sources is an effective way to characterize urban soundscapes pleasantness, and 2) physical descriptors computed for specific sound sources better characterize the appraisal than global descriptors.
  \end{abstract}

  \begin{keyword}
  %% keywords here, in the form: keyword \sep keyword
 cognitive psychology \sep soundscape assessment
  %% PACS codes here, in the form: \PACS code \sep code

@@

One of the main goal of soundscape studies is to identify which components of the soundscape influence human perception \cite{aletta2016soundscape}. For example, in order to study the link between an environment (urban sound scene) and the induced human sensation (calmness), one mean is to objectify the mental representation of a specific sound scene, a calm urban sound scene. Several means have been considered in the literature to do so.

@@

First, a subject can be asked to describe a given sound following a given perceptual scale (\eg~calmness, pleasantness) \cite{axelsson2005soundscape,davies2013perception,cain2013development}. The amount of information that can be gathered strongly depends on the very nature of the available stimuli, be they recorded sound scenes as part of a within laboratory experiment, or \emph{in situ}. That being said, those stimuli respectively are or can be recorded and analysed. It thus allows the experimenter to gather a coarse grain physical description of the stimuli.

@@

Second, a subject can be asked to describe a given sound environment \cite{guastavino2006ideal, dubois2006cognitive}. A large amount of quantitative and semantic information is then collected about the representation of this type of sound environment that the subject has. Unfortunately, without any reference to sound data, this representation can hardly be characterized physically.

@@

We propose in this paper to consider the use of a soundscape simulator that the subject can use to objectify her representation of a given sound environment. We believe that the use of such a device allows us to gain the benefits of the two above cited approaches. As the subject is asked to produces audio data (the signal of the simulated scene), it allows the experimenter to study a precise modal version of the mental representation of the subject that is both characterized semantically and physically, respectively in terms of the sound sources and their levels.

@@

For several reasons that are detailed in this paper, the use of a soundscape simulator such as the one proposed in this article can lead to interesting outcomes as it allows the experimenter to separately study  the influence of the sounds sources that compose a sound environment of interest. With such a material, not only the type of sound sources is indeed available for study, but also the exact level and audio waveform for each source together with the structural characteristics of the scene, that is the temporal distribution of the events.

@@

@@

To the best of our knowledge, only Bruce~\cite{bruce2009development,bruce2014effects} considered the use of a simulator to question soundscape perception. They propose a tool that allows the user to modify a given soundscape by adding or removing specific sound sources, by changing the acoustic level of the sources as well as their spatial location. The authors shows that the addition or removal of the sources globally follows more social or semantic matters than their acoustical characteristics. A lack of diversity is nonetheless mentioned by the authors that limit the strenght of the outcomes given in this study.

@@

The remaining of the paper is organized as follows: the soundscape simulator \emph{simScene} is introduced in Section \ref{sec:simulator}. The experiments  1.a \emph{simulation} and  1.b \emph{pleasantness evaluation} are presented in Section \ref{sec:simulation}. Experiment 2 \emph{pleasantness evaluation after modification of the scenes} is presented in Section \ref{sec:modification}. Conclusions and discussion about future work follow in Section \ref{sec:conclusion}. Experiment 1.a is based on a pilot study described in \cite{lafay2014new}.

@@

\emph{Simscene} is an online digital audio tool whose first version have been developed as part of the HOULE project\footnote{\emph{Projet HOULE} : \url{houle.ircam.fr}}. It has been designed to run on the popular web browsers \emph{Chrome} and \emph{Firefox}. It is fully written in javascript using the \emph{angular.js} library\footnote{\emph{angular.js} : \url{angularjs.org}} and the \emph{web-audio} standard that allows the manipulation of digita audio data within the browser\footnote{\emph{web-audio}: \url{www.w3.org/TR/webaudio}}. The interface for selecting the sound sources (\cf~Section~\ref{sec:simscene_ssf}) uses the popular \emph{D3.js} \cite{d32011} visualization library.

@@

\emph{Simscene} is designed as a simplified audio sequencer, with sequencing parameters specifically chosen for the generation of realistic soundscapes. To do so, the user first selects a sound source using a non verbal selection interface presented in Section~\ref{sec:simscene_ssf}. A track is then created for this sound source within the simulator interface. The user can then manipulate some parameters detailed in Section~\ref{sec:simscene_parametre} to control the time and magnitude distribution of the occurrences of the sounds source. Text fields are also available for the user to 1) name each track, 2) name the entire scene, 3) to provide some comments about the simulated scene.

@@

In order to provide the user with a sound database that is well organized and covers as much as possible the variety of sound sources that are available in urban areas, a typology of urban sounds is first defined.

@@

@@

The literature usually distinguish between:
\begin{itemize}
\item {a sound event}: an isolated sound, sparsely distributed whose acoustical characteristics may change with respect to time
\item {a sound texture}: an isolated sound of long duration whose acoustical characteristics are stable with respect to time \cite{saint1995classification}.
\end{itemize}

The simulation tool \emph{Simscene} follows this distinction and consider two distinct sound databases: one with the classes of events only and the other with classes of textures only. These two types also have specific scheduling procedures during the simulation process.

@@

@@

According to the Auditory Scene Analysis (ASA) theory, the brain processes separately events emerging from different sources of a sound environment \cite{bregman1994auditory}. Several auditory streams are created, one for each sequence of events emitted by the same source \cite{carlyon2004brain}. In this case, the source is a perceptual entity as events that are not isolated are grouped into a stream.

@@

Maffiolo \cite{maffiolo_caracterisation_1999} distinguish two distinct categorization processes, that differentiate themselves with the capacity of the listener to identify sound events. She shows that those processes leads to two abstract cognitive category respectively termed "the event sequences" and "the amorphous sequences".

@@

The event sequences are composed of salient events that can easily be recognized such as \emph{car start} or \emph{male speech}. They arise from a descriptive analysis based on the identification of the sound sources. On contrary, the amorphous sequences are sound environments where distinct events can hardly be found, such as \emph{traffic hubhub}. They result from a holistic analysis based on global acoustic features.

@@

\begin{figure}[t]
        \myfloatalign
        \includegraphics[width=.8\linewidth]{gfx/ch_4/3.png}
       \caption{Hierachical organisation of the isolated sounds used of the simulation.}\label{fig:orgDb}
\end{figure}

@@

We term sample, a recording of an isolated sound, be it an event or a texture. Each sound class is implemented as a collection of samples perceived equivalently.

@@

The sound classes are organized hierarchically (\cf~Figure~\ref{fig:orgDb}) according to a structure that is close to the one of the vertical axis of the categorical organization proposed by E. Rosch \cite{rosch1978cognition}. The lower the level of abstraction, the more precise the description of the class and the more perceptually similar the sound sources is. For classes with high level of abstraction that have sub classes, its collection of samples is the union of the collections of the sub classes.

@@

Accounting for the perceptual matters detailed previously, two taxonomies are built, one for event sounds and one for texture sounds. Four levels of abstraction are considered from the more generic one (level 0), to the more specific classes (level 3), leading to a taxonomy close to the one used in \cite{Salamon14}. Only three levels of abstraction are considered for the texture sounds.

@@

All samples are normalized to the same $RMS$ level of $-12$ $dB$ FS, \ie relative to Full Scale. In our case, the full scale level is of 1 Volt.

@@

By a volountary design choice, the simulation tool do not allow the user to interact and control a specific sample. Interaction is done at the track level, a track being a sequence of samples. Several parameters are available to the subject to control the track:

@@

\begin{itemize}
\item \emph{level} ($dB$): the levels are  randomly sampled according to a normal distribution parametrized by the subject in terms of mean value and variance;
\item \emph{inter-onset spacing} (second): for the event's track only, the spacings are randomly sampled according to a normal distribution parametrized by the subject in terms of mean value and variance;
\item \emph{start and end time} (second): the subject set the first and last occurrence of a sample of the selected class.
\end{itemize}

@@

\begin{itemize}
\item \emph{event's fades} (seconds): for the event's track only, the subject can set a fade in / fade out duration applied to each samples;
\item \emph{global fades} (seconds): the subject set the duration of the fade in / fade out at the beginning and the end of the track.
\end{itemize}

@@

@@

Once the typology and the set of sounds are available, an important design issue is the need of a suitable way to display the sound dataset to the user. Most browsing tools are based on keyword indexing. However, for sensory experiment that study the objectivation of the mental representation of the subject, this mean is problematic as the availability of a verbal description of the sound can influence the choice of the subject and potenitally induce biases in the analysis conclusions. Indeed, the subject may select \textit{a priori} sounds referenced as belonging to \emph{park} to build a \emph{calm} soundscape.

@@

Therefore, the selection interface considered in this study is text free and composed of spatially organized circles on a plane.

@@

\begin{figure}[t]
        \myfloatalign
        \subfloat[]
        {\includegraphics[width=.8\columnwidth]{gfx/ch_4/SSF}\label{fig:ssf}} \par
        \subfloat[]
        {\includegraphics[width=\columnwidth]{gfx/ch_4/simscene}\label{fig:simscene}}
       \caption{Graphical interfaces for the selection of sound classes (a) and their sequencing (b) of the \emph{simScene} tool.}
\end{figure}

@@

As shown on Figure~\ref{fig:simscene}, the simulation interface displays a schematic of the scene under creation. Each track is represented as an horizontal band with a temporal axis. Each sample of this track is displayed by a rectangle whose height is proportional to the amplitude of the sample. For event's tracks, the horizontal spacing between those rectangles is a function of the time delay between their onsets. For tracks of texture, a unique rectangle is displayed as this kind of sounds do not allow spacing with silence. As actual amplitude and spacing values are drawn from random variables, each time the subject changes the value of a parameter, the location and height of the rectangles are updated to reflect the changes in the sequencing of the samples. The subject can listen to the resulting sound scene at any time.

@@

% L'objectif est d'étudier les influences spécifiques des différentes sources sonores qui composent les environnements urbains sur la perception de l'agrément, en utilisant la simulation. Pour ce faire, nous planifions nos deux premières expériences (\cf~Figure~\ref{fig:xp1_2}):
%
% \begin{itemize}
% \item \emph{expérience 1.a de simulation}: au cours de cette expérience, les sujets doivent simuler des environnements sonores urbains, en utilisant l'outil de simulation \emph{Simscene} (\cf~Section~\ref{sec:simscene}). Chacun compose deux environnements sonores, le premier idéal/agréable et le deuxième non-idéal/désagréable.
%
% \item \emph{expérience 1.b d'évaluation}: à l'issue de la simulation, nous n'avons, de fait, qu'une connaissance binaire des propriétés affectives des scènes simulées: idéale (i) et non-idéale (ni). Cette seconde étape a pour but d'affiner notre connaissance sur l'agrément. Pour ce faire, nous demandons à un deuxième groupe de sujets d'évaluer, à partir d'une échelle sémantique, l'agrément de chacune des scènes simulées. L'expérience d'évaluation sert deux buts:
%
% \begin{enumerate}
% \item évaluer l'influence respective des différentes sources sur l'agrément, pour chaque type d'environnement (i ou ni);
% \item détecter la présence de cas extrêmes ou ambigus (\emph{outlier}) dans les scènes simulées. Pour le reste de notre étude, les qualités hédoniques imposées (i et ni) servent de référence, de vérité terrain. Il nous faut donc garantir qu'il n'y ait pas d’ambiguïté entre les cas extrêmes des i- et ni-scènes, \ie~que la note d'agrément la plus basse des i-scènes reste supérieure à la note la plus haute des ni-scènes.
% \end{enumerate}
%
% \end{itemize}
%
% Notre analyse s'appuie sur les données produites par les deux expériences.

Experiment 1 aims at using a simulation paradigm to investigate the specific influences of the various sound sources constituting urban soundscapes on the perceived pleasantness. For that, the two first experiments are planned as follows (\cf~Figure~\ref{fig:xp1_2}):
\begin{itemize}
\item \emph{experiment 1.a (simulation)}: during this experiment, participants are asked to create urban sound environments, using the simulation tool \emph{Simscene}, see Section~\ref{sec:simulator}. Each of them creates two environments: one ideal / pleasant and the other one, non ideal / unpleasant.
\item \emph{experiment 2.a (evaluation)}: after the simulation phase, only a binary information on the pleasantness property of the respective scenes: respectively ideal (i) or non ideal (ni). Furthermore, this information is given by the creator of the scene. The second experimental step aims at investigating more deeply and more broadly our knowledge on pleasantness. For that, a second group of participants are asked to evaluate the pleasantness of each simulated scene, on a semantic scale. This experiment has two goals:
\begin{enumerate}
\item evaluate the respective influence on the pleasantness of the various sources composing the scenes (i or ni)
\item detect the presence of \emph{outliers} or ambiguous samples. For the remainder of the study, the defined hedonic properties (i and ni) are used as references. We thus need to ensure beforehand that no ambiguity exists between extreme cases of i and ni scenes; i.e that the lower i-scene is really above the upper ni-scene,  in terms of perceived pleasantness.
\end{enumerate}
\end{itemize}

Our analysis is based on the data collected by these two experiments: 1.a and 1.b.

\begin{figure}[t]
        \myfloatalign
        \includegraphics[width=\linewidth]{gfx/ch_5/5.png}
        \caption{Experimental protocol of the simulation experiment (1.a) and the pleasantness evaluation experiment (1.b)}\label{fig:xp1_2}
\end{figure}

@@

@@

% Les sujets doivent simuler deux environnements sonores urbains, chacune des scènes devant durer 1 minute. Ils doivent se conformer aux consignes suivantes:
%
% \begin{itemize}
% \item première simulation: simuler un paysage sonore \textbf{urbain plausible} qui, selon vous, est idéal (où vous aimeriez vivre);
% \item deuxième simulation: simuler un paysage sonore \textbf{urbain plausible} qui, selon vous, est non-idéal (où vous n'aimeriez pas vivre).
% \end{itemize}
%
% Tous commencent par simuler l'environnement idéal. Ils ne prennent connaissance de la deuxième consigne qu'à la fin de la première simulation.
%
% Les sujets sont totalement libres dans le choix des sons, et des paramètres. Ils doivent cependant se soumettre à deux contraintes:
%
% \begin{itemize}
% \item le sujet doit prendre le point de vue d’un auditeur fixe;
% \item le paysage sonore doit être réaliste, au sens de physiquement plausible. Le sujet peut très bien réunir 10 chiens dans un paysage sonore. Il ne peut en introduire 1 qui aboierait toutes les 10 millisecondes.
% \end{itemize}
%
% Ces contraintes font partie de la consigne. Aucun contrôle n'est fait \emph{a priori} dans l'interface de simulation.

The participants have to simulate two urban sound environments of one minute each. They follow the instructions below:
\begin{itemize}
\item  first simulation: create a \textbf{plausible urban soundscape} which is ideal, according to you (where you would like to live);
\item second simulation: create a \textbf{plausible urban soundscape} which is non ideal, according to you (where you would not like to live);
\end{itemize}

All the participants start by designing the ideal environment; they read the second set of instruction at the end of the first experiment. Participants are completely free of their choice concerning sounds and synthesis parameters. They must nevertheless respect the two following constraints :
\begin{itemize}
\item the listening point of view is the one of a fixed listener;
\item the soundscape must be realistic, i.e. physically plausible.
\end{itemize}
For instance, the participant can insert ten dogs in the soundscape but she cannot insert one dog barking every 10 milliseconds.
These constraints are notified in the instructions; no control is done \emph{a priori} in the simulation user interface.
